---
title: "R Penguins"
output: html_document
date: "2025-06-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The first thing we need is <b> data</b>. For this project, we will be using R's penguins_raw dataset.

In order to access this dataset, we must first download and install the R package "palmerpenguins." 
A <b> package </b> is a set of functions, code, and data made by developers that other developers are free to download and use in their code. 
Once we've installed the package, we then need to load it with the library() function.
This simply means that we are telling our file-- in this case, our R environment-- to use the package in the project. 

Think of it like this. You are the professor, and R is the student in a class. Telling R to install the package is like telling R to buy a textbook. Calling the library() function is telling the student to open and read the book so they can use it in class. 
```{r}
install.packages("remotes", repos = "https://cloud.r-project.org/")
remotes::install_github("allisonhorst/palmerpenguins")
```
```{r}
library("palmerpenguins") 
```

Great! Now that we've run this code, the library is ready for us to use.

Now, let's get a feel for the data. To do this, we will run the head() function.

```{r}
head(penguins_raw)
```

Try it now, what do you see? 

This type of table is called a <b> dataframe </b>. A dataframe is not unlike a table or an Excel spreadsheet. It is an easy way to organize the data in a two-dimensional structure. 
The head() function returns the first 6 rows of a dataframe. 

What information do you get from the head? What information do you not get? 

<!-- Interactive box or something to brainstorm... it's like you can write bullet points--> 

The first thing you should notice that you get from the head is the names of the columns of our dataframes, as well as their datatypes. 

Another helpful function is the summary() function. 
Let's try it out on our dataset:

```{r}
summary(penguins_raw)
```

This function helps us get a feel for our data in a different way. What information do you get here that you don't get from head()? 



<!-- https://www.twilio.com/en-us/blog/developers/best-practices/dataset-questions-to-ask 
This describes different things to do with a dataset. We will work with these questions from the framework.-->