---
title: "Intro to R: Programming Basics"
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
tutorial_options(exercise.eval = FALSE)
gradethis::gradethis_setup()
```

```{r prepare-variables}
mary_age <- 22L
```

# Variables

What are variables?

A variable is a named container for data. 

<!-- Uhh home canners write the name of the stuff and the date that's being canned. Think of it like that. You're telling the computer "name something and keep the information about it" so you don't have to --> 

Meet Mary. 

<!-- doodle or something cute -->

We want to store some information about her, that we want to be able to easily access. 

<!-- Doodle of Mary saying she's 72 inches tall --> 
Say Marys's age is 60, lets make a variable to store that information and be able to access it later.
In R, we use the symbol <code> <- </code> to create variables. 
The fancy computing term for this, <code> <- </code> is <b> assignment operator </b>. 

```{r, eval=TRUE, echo = TRUE}
mary_height <- 72
```

Let's double check that we saved her age correctly. R has a built in <code> print </code> function that will, as the name suggests, print a value. Click "Run Code" to see it in action.

```{r, print_var, exercise=TRUE, exercise.setup = "prepare-variables"}
print(mary_height)
```

Great! Since it printed 72, we know that <code> mary_age </code> is saved correctly. 

How tall would she be if she was 5 inches? You can probably do this in your head, but let's see if the computer can do it. 

```{r, first_var, exercise=TRUE, exercise.setup = "prepare-variables", eval=TRUE, echo= TRUE}
print(mary_height + 5)
```

Notice that it automatically computed the number [CHANGE THE WORDING] based on the value of our variable <code> mary_height </code>. 
This is the advantage of using variables: it allows our code to be dynamic. 

Why is this dynamicness (dynamicity? idk) important? 

Well, Mary has a confession to make. 

<!-- Mary doodle, she lied about her height! She's actually 60 inches tall. -->

Oh boy, how do we handle this?

We can simply change the variable value. This is called <b> what's the vocab. reassigning? </b>. 

```{r, update_var, exercise=TRUE}
mary_age <- 60
print(mary_age)
```

Copy and paste the same code from the earlier block, where we computed Mary's height adjusted by five inches 

<!-- Maybe add like a thing of "copy/paste the same code from the earlier block. Look! We didn't change our code, but mary_age + 5 is updated. That's one of the advantages of using variables. You can easily change the value without having to manually go back and mess with blah blah blah code. 
I can't figure out how to add the code block. --> 

This is a small example but it's illustrative of something important: by using variables, you can quickly adjust your code without having to go through and manually rename everything. 

<!-- Ella make this more eloquent --> 

Variables can hold other types of information than just numbers! In the coming sections we will explore several other types of information we can store in variables.

<!-- Maybe add an exercise here, assign your own variable for Mary's friend Miguel. How tall is he? Print it to make sure it's correct. Just make sure they create a float/whatever it's called and print it.--> 

## Floats and Numerics 

<!-- Beginnings of data types. What is a float? 

I just moved this here, you can edit content how you like to make it make sense--> 

<!-- introduction to numeric operations etc. --> 

Integers are whole numbers, what if we want to store something that isn't a whole number, like a measurement? In programming, this is a distinct type of data called a float.

The main different between a float and an integer is that a float stores a number with a decimal point, rather than a whole number.

<!-- R defaults to floats when you assign a number. Try to think of some variables that would work well as float! height, weight, temperature, angles, volume --> 

In R, these are also referred to as a "numeric"

Try storing a decimal to `my_float` and watch the output to confirm it worked. Feel free to use the "hint" button if you need more help.

```{r float-example, exercise=TRUE}
my_float <- 
print(my_float)
```

::: {#float-example-hint}
**Hint:** Try typing numbers like 3.14, 9.8, or 6.022 after the "\<-"
:::

```{r float-example-check}
grade_result(
  pass_if(~identical(typeof(my_float), "double") & (my_float %% 1 != 0)),
  fail_if(~identical(typeof(my_float), "double") & !(my_float %% 1 != 0), message="R techinically considers your response a float because there isn't an \"L\" after it, but it is best to try again with a decimal value for the sake of the exercise.")
)
```

<!-- Maybe add an exercise just saying. R has built in operations that you know from math class. 
+, -, *, / 

Introduce modulo? is that a can of worms. Operators can also be a separate lesson, what are your thoughts --> 


## Integers

Integers store our countable numbers. 

<!-- I think this bit should introduce age. --> 

<!-- Justification: It's unusual to have ages as a float. We don't say I'm 32.64829 years old; you say I'm 32. --> 
This is the type of variable we used to store Mary's age. Try creating one yourself above the `print()` function. Name it `my_integer` and set it to whatever you like! Once you're done, click "Submit Answer" to check your work.

```{r, int-example, exercise=TRUE}

print(my_integer)
```

```{r int-example-check}
grade_result(
  pass_if(~is.integer(my_integer)),
  fail_if(~is.double(my_integer), message = "You are either missing something after your number, or you aren't typing in a whole number. Check the above examples.")
)
```

Integers can be added. Here are a few ways to do so.

```{r, int-add, exercise=TRUE}
  int1 <- 8L
  my_integer_2 <- 4L

  print(int1 + 4L)
  print(int1 + my_integer_2)
  
```

Experiment with the exercise above and confirm integer addition works as you expect.

## Strings

Strings can store words and sentences, but they'll primary be used for words in R.

As an example, lets store a person's name.

```{r, string-example, exercise=TRUE}
my_name <- "Miguel"
print(my_name)
```

See if you can change the code below to print "My name is Miguel"

```{r, string-sentence, exercise=TRUE}
my_name <- ""
print(my_name)
```

```{r string-sentence-check}
grade_result(
  pass_if(~identical(.result, "My name is Miguel"))
)
```

## Booleans

<!-- Lead with the example. We're building our profile on Miguel. We know his age (float), his XYZ (int), his name (string). Now we want to store a value of whether he likes pizza or not. This can have two values TRUE or FALSE. that's our window into boolean--> 

Booleans can only store one of two values, true or false. This is great for data that can only have two possibilities, or to represent a "yes" or "no".

<!--Kind of a silly example, I plan to change it-->

What if we want to represent whether or not Miguel likes pizza?

```{r bool-example, exercise=TRUE}
miguel_likes_pizza <- TRUE
print(miguel_likes_pizza)
```

How would you represent Miguel not liking pizza?

```{r bool-exercise, exercise=TRUE}
miguel_likes_pizza <-
print(miguel_likes_pizza)
```

```{r bool-exercise-check}
grade_result(
  pass_if(~!(miguel_likes_pizza)),
  fail_if(~typeof(miguel_likes_pizza) != "logical", message="Remember that booleans can only hold \"TRUE\" or \"FALSE\"")
)

```

## Other Datatypes

There are more complicated datatypes, such as more complex numbers or time, but this is too in the weeds for right now, so don't worry about it. Just keep in mind that our list isn't exhaustive.

# Functions

Functions are reusable chunks of code that can be built-in to our environment, or be made by us.

<!-- If variables are nouns, functions are verbs. Maybe we can do something with the person profile!

1. Function age_in_five_years -> return age + 5
2. Function is_old -> for conditional 
3. We don't need to cover loops do we?
Like we want a function is_old() (NOT A GREAT IDEA JUST WHAT'S COMING TO MIND). If age geq 100, return TRUE. else return false. 

Maybe that's too much to introduce a conditional. --> 

The purpose of a function is to label a repeated task (chunk of code), to make a program neater. It is similar to a variable but stores lines of code rather than a single data value.

We've already used a function to show how variables work, `print()`. As you can probably tell, `print()` takes a variable in its parenthesis and "prints" it to the screen. The name for variables that go in a functions parenthesis are "arguments". Therefore, we can say that \`print() function takes one argument and prints it to the screen.

`times_two()` is a simple multiplication function that takes one parameter, and multiplies it by 2. See if you can get it to multiply 3 by 2.

```{r, function-example, exercise=TRUE, exercise.lines = 5}
times_two <- function(num) {
  return(num * 2)
}
```

```{r, function-example-solution}
times_two <- function(num) {
  return(num * 2)
}
times_two(3)
```

```{r, function-example-code-check}
grade_code()
```
