---
title: "Intro to R: R Basics"
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
tutorial_options(exercise.eval = FALSE)
gradethis::gradethis_setup()
```

R is a weird language. 

We've built out a profile for Miguel and Sarah (or whatever it is). 

However, everything is stored in separate variables. If only there was a way to store things together! 

## Vectors

Miguel wants to do a study on the how many slices of pizza he eats each time he has pizza. 

R is very data friendly. We can store this data as a <b> vector </b>.
 
```{r}
miguel_pizza_consumption_data <- c(12, 6, 2, 3, 1, 4, 5)
```

We use vectors when we want to store lots of entries of the same type (see the previous lesson for data types).

```{r}
miguel_pizza_enjoyment <- c(FALSE, TRUE, TRUE, TRUE, FALSE, TRUE)
```

Miguel extends his study by a day. 

```{r}
new_pizza_consumption_data <- append(miguel_pizza_consumption_data, 2)
```

Uh oh! Miguel marked his notebook wrong.
```{r}
miguel_pizza_enjoyment <- append(miguel_pizza_enjoyment, 2)
```

what happened to the data type? 

```{r}
print(miguel_pizza_enjoyment)
```

Let's fix it

```{r}
# Change the last entry 
# Do as.logic()
```

as. is a helpful family of functions. 

Say we want to see the his results on the third day. 

```{r}
miguel_pizza_consumption_data[3]
```

He ate two pieces of pizza!


## Other Trivia

rep(), : sequencing

More just worth mentioning. Not to worry about.

## Lists

We want more than one type of datatype. 

List. 

```{r}
miguels_favorite_things <- list("pizza", 2, "dog", TRUE)
```

Has similar things as vectors for indexing, changing values. 

## Matrices 

Worth mentioning, helpful for linear algebra. I don't think we need to push this too much.

## Data frame

Make a dataframe of the profiles. 

columns are "name" "age" "likes_pizza" 
Helpful for data, most similar to excel spreadsheets

We will work with dataframes more. 

Each column is like a vector and contains same data type

## Factors

What is a factor? Why important? 

Start with a vector/list example. 
Then give the dataframe. 

Go over levels().
Compare length() on a factor vs a list.
