---
title: "Intro to R: Loading Your Data"
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
tutorial_options(exercise.eval = FALSE)
gradethis::gradethis_setup()
```

# Introduction

For this project, we will be working with morphological data collected on the Black Oystercatcher, or <i>Haematopus bachmani. </i>

First, we need to get the data onto our computer and into R so that we can use it. 

Right now, the data is stored in a .csv file. If you've ever worked with a spreadsheet in Excel, Numbers, or a similar app, then you've seen something similar to a .csv before. 

.csv stands for "comma-separated values." .csv files in practice work very similar to tables like you see in Excel. 
<!-- Make this link nice and clickable- maybe even a big button -->
Take a look at a raw csv file here:
<button onclick="window.location.href='https://raw.githubusercontent.com/ellaroxanne/r_first_data_project/refs/heads/main/data_and_metadata/BLOY_morphology_AlaskaBC.csv';">BLOY Morphology Data</button>

## Question 1 
Why do you think it's called a comma-separated value file? 

<!-- Text box to answer-->

R, and many other computer programs, does a good job at organizing a csv file into a dataframe, which is similar to an Excel spreadsheet or other tabular (table-like) form of organizing data. We just need the computer to read it-- to turn it from this raw file into a nice table. 

What better function than <code> read.csv() </code>!

## Your Task: 
Write a piece of code in R that uses the <code> read.csv() </code> function to save a variable named <code> raw_data </code>. 

### Step 1: Naming a variable

A variable is essentially giving a name to a piece of information you want to store. This is called initialization. 
Let's start by making a variable called <code> my_favorite_number </code>. 

To do so in R, you start with your variable name, then an ``arrow'' <code> <- </code>, and then the value you want it to be. 

So, a code for my favorite number would be:

```{r echo = TRUE}
ellas_favorite_number <- 2
```


Try running your own, and see what happens!

<!-- Space for user to try one -->

Now, let's double-check that variable saved correctly. 
A straightforward waya to do this with simple variables is the <code> print </code> function.

```{r}
print(ellas_favorite_number)
```
(Hint: If this isn't working, make sure you've run your previous block, where ellas_favorite_number is initialized.)

Return to the excerise and try it with yours!

<!-- Space for user to write "print(my_favorite_number)"-->

<!-- Exercise with a string?-->

Congratulations! You've made your first variable. 

### Step 2: Enter .csv

Our eventual goal is to store the information contained in our .csv as a variable called <code> raw_data </code>.

But how can we extract this information? 

Let's start by trying to make a variable with the link we have, https://raw.githubusercontent.com/ellaroxanne/r_first_data_project/refs/heads/main/BLOY_morphology_AlaskaBC.csv. 

<!-- Have them copy this code, that's down below-->

```{r eval=FALSE}
first_try <- "https://raw.githubusercontent.com/ellaroxanne/r_first_data_project/refs/heads/main/BLOY_morphology_AlaskaBC.csv"
```

Why the quotation marks? 

When you're trying to type words, links, or basically anything other than numbers or True/False values (called Booleans), you need to write them as a <b> string </b>. 

The quotation marks around them essentially tell the computer, "this is a word." They act as a container for the information that the computer will store. 

Let's use our <code> print </code> trick to see that the variable saved the way we wanted.

```{r eval=FALSE}
print(first_try)
```

Uh oh! 

Computers are very literal, so when we tell it to save our .csv link as a variable, that's what it saves: the link. Not the information inside. How do we unpack it? 

### read.csv()

Because .csv files are so important and ubiquitous in data science, R has a built in function that allows you to easily extract the information from a .csv and transform it into a table, which is much easier both to visualize and to use. 

This function is called <code> read.csv() </code>. 
As an argument, it takes the .csv file that you want to turn into a table, or DataFrame. 

Use the <code> read.csv() </code> function to create a variable named raw_data using the csv link we have. 

<!-- This is our answer code; I want an interactive code box where they can try. The answer is revealed on the end or if they need a hint-->

```{r eval=FALSE}
raw_data <- read.csv("https://raw.githubusercontent.com/ellaroxanne/r_first_data_project/refs/heads/main/BLOY_morphology_AlaskaBC.csv")
```

Alright! It looks like we've loaded our data, let's try our <code>print </code> trick to take a look:

<!-- Have the user type this 
I have it commented out just bc it's huge in the file -->
```{r eval=FALSE}
# print(raw_data)
```

Woah! That's huge!

Since we're just getting started, we don't need all that information. In fact, it may even be too much information; it's totally normal to be overwhelmed by this!

In fact, it's so normal, that R developers created a specific function to help you get a basic idea of your data, <i> without </i> overwhelming you, or your computer, with code. 

## Basic Functions: Dataframes, head(), colnames()

The first function we'll take a look at is the <code> head </code> function. 

Let's start by getting a basic feel for the data with the head() function. Try it now:

```{r eval=FALSE}
head(raw_data)
```

### Question: What does the <code> head() </code> function do? 
<!-- Answer box; eventual answer will just be "Displays the first six rows". Perhaps we can have a way for the computer to check, or self check-->

### Question: What information do we get from the <code> head() </code> function? What information don't we get? 
<!-- Brainstorming space as well. --> 

<code> head() </code> gives us a nice way to get a feel for our data without printing all of it. 
It shows us all of the variables measured, and gives us an idea of what they may look like. 

<!-- Note to self: perhaps there should be some explanation between variable (c.s.) and variable (stats) -->


Another helpful function when getting to know your data is <code> colnames() </code>. Try it now.

```{r eval=FALSE}
colnames(raw_data)
```

### Question: What information do you get from <code> colnames() </code>? 

<!-- Text box. Eventual answer: a list of each variable that was collected for this data set--> 

### Food for Thought: When do you use colnames() in addition to head()? 

<!-- brainstorming box--> 
