---
title: "02-Metadata"
output: html_document
date: "2025-07-10"
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
raw_data <- read.csv("https://raw.githubusercontent.com/ellaroxanne/r_first_data_project/refs/heads/main/BLOY_morphology_AlaskaBC.csv")
```
Alright! We've loaded our data, and gotten a basic idea of what it looks like. Does that mean we're ready to start our studies? 

Not quite. 

Sure, you could technically jump striaght into crunching numbers, reporting findings, etc. However-- that would not make you a very responsible data scientist. 

Before working with data, you need to know your data. Know the informatin about the data. 

This is so important, it has its own name: <b> metadata </b>. Basically, data about data.  

Relevant links 
(Will compile)

We already know that this data is on the Black Oystercatcher. 
## Question: What is the Black Oystercatcher? 
<!-- Progressive hints; use this link, on this page, etc.-->

The most basic part of this is to understand what your data means. Let's return to the <code>colnames()</code> function we learned about earlier:
```{r}
colnames(raw_data)
```
Clearly, this data shows information on various features of Black Oystercatcher specimens, collected as a part of this study.

## Question: What do each of the column names mean? 

<!-- Perhaps a table or something, with each column name on the left and then a space to type your answer on the right. Or, easy mode like a click and drag matching. --> 
<!-- Or start with the easy ones, offer just a helping guide. Latitude, longitude, date, mass, etc.--> 

Before we actually get to working with the data, it's best practice as a data scientist to consider what kind of data you are working with. Numbers don't appear out of a vacuum. 

Here are some relevant links to help you answer the following questions: 

## How was the data compiled?
 [text box to answer]
 
## Is the data accurate?
[text box to answer]

Great! Now that we've affirmed the quality of our data, let's get down to what we're really interested in: the data itself.

We've already done the head() function, which gives us a general idea of what kind of variables (columns) are being measured. A <b> variable </b> is a quantity or quality of each data point that is measured. 

Especially with large data files such as this one, it can be difficult to fully wrap your head around the head function. Here, we can use the colnames() function.

<!-- I want the user to be able to type this function in -->

```{r}
colnames(raw_data)
```


Some of these variables, such as "ParkName" or "ScientificName" are fairly intuitive. 
Some are not. 

<!-- Exercise: Identify in 1-2 sentences what each variable does-->

Now that we know the data, we'll do some summary statistics in the next module.




