---
title: "Intro to R: Basic Summary Statistics"
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
tutorial_options(exercise.eval = FALSE)
gradethis::gradethis_setup()
```

```{r}
raw_data <- read.csv("https://raw.githubusercontent.com/ellaroxanne/r_first_data_project/refs/heads/main/data_and_metadata/BLOY_morphology_AlaskaBC.csv")
```

Now that we vaguely know what our data's about, let's get a little more used to what it contains. 
Yay! We finally get to work with numbers!

[add a cute gif of someone celebrating]

Let's make sure our data is clean. First, we'll have the is.na() function. This checks for if there are any values missing. 

```{r}
is.na(raw_data)
```

Whoa! That's a lot of information. It's hard, and easy to miss if there's a TRUE in there-- meaning, there's a value that is indeed missing. 

```{r}
na_dataframe <- is.na(raw_data)
sum(na_dataframe)
```

Aha! There's 85 missing values. Great!

...

Where are they? 

We can do this by columns. R has a built-in function called <code> colSums </code> that is made for this purpose. It basically sums by column.

```{r}
colSums(na_dataframe)
```

This process is called <b> data cleaning </b>. 